---
title: "01_preprocess"
decription: "Generating variables, cleaning and preprocessing for modelling and plotting"
project: "Vocatives"
author: "Marton Soskuthy & Timo Roettger"
date: "7/24/2019"
contacts: "timo.b.roettger@gmail.com, marton.soskuthy@ubc.ca"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

Loading libraries & raw data.

```{r}
library(tidyverse)
library(purrr)

v_orig <- read_csv('../raw_data/vocatives.csv')

```

Some basic formatting, e.g. numeric 0-1 to logical. 

```{r}
colnames(v_orig) <- gsub(" ", "_", tolower(colnames(v_orig)))
v <- v_orig %>%
  mutate(default=!is.na(default),
         prosody_clean=!is.na(prosody_clean),
         prosody_stress=!is.na(prosody_stress),
         prosody_tone=!is.na(prosody_tone),
         prosody_vowel_lengthening=!is.na(prosody_vowel_lengthening),
         morphological=dplyr::recode(morphological, yes='affix'),
         noun_class=replace_na(noun_class, ""),
         number=replace_na(number, ""),
         conditioning=replace_na(conditioning, ""),
         form_add=replace_na(form_add, ""),
         form_replace=replace_na(form_replace, ""),
         form_delete=replace_na(form_delete, ""),
         need_checking=!is.na(need_checking),
         exclude=!is.na(exclude)) %>%
  filter(!need_checking, !exclude) %>%
  dplyr::select(-case_comment, 
         -epenthesis,
         -intonation,
         -`function`,
         -optional,
         -reference,
         -notes,
         -quotes,
         -need_checking,
         -exclude)
```

## Case column

```{r}
v <- v %>%
  mutate(case_broad=dplyr::recode(case,
                                  accusative="acc-like",
                                  nominative="nom-like",
                                  vocative="vocative",
                                  absolutive="nom-like",
                                  ergative="acc-like",
                                  oblique="acc-like",
                                  `accusative/dative`="acc-like",
                                  `object marker`="acc-like",
                                  agentive="nom-like",
                                  `"nominative"`="nom-like",
                                  `non-nominative`="acc-like",
                                  `anti-agentive`="acc-like",
                                  `absolutive-modalis (indefinite object)`="acc-like",
                                  `relative`="acc-like",
                                  `subject case`="nom-like",
                                  `objective`="acc-like",
                                  `direct object`="acc-like"))
```

## form_add-based variables

Creating some cleaner versions of form-add.

- single character for each consonant
- separate prefix vs suffix 

```{r}
conversion_table <- list(
  `ts[']*`="K",
  `tʃ[']*`="C",
  `dʒ`="J",
  `d´`="d",
  `ph`="p"
)

# for converting digraphs to single characters
transform_to_single <- function (str_vector, conversion_table) {
  for (key in names(conversion_table)) {
    str_vector <- gsub(key, conversion_table[[key]], str_vector)
  }
  return(str_vector)
}

v <- v %>%
  mutate(form_add_s=transform_to_single(form_add, conversion_table),
         form_add_s=dplyr::recode(form_add_s, zero=""),
         form_add_prefix=unlist(map(str_split(form_add_s, "-+"), function (x) x[1])),
         form_add_suffix=unlist(map(str_split(form_add_s, "-+"), function (x) x[length(x)])),
         form_add_all=gsub("[-]+", "", form_add_s))
```

Let's generate some form_add-based variables. The LIST:

- vowel vs. no vowel (1vs0)
- does it end in a vowel or not (1vs0) 
- if C is it sonorant?
- Presence of voiceless obstruent (1vs0)
- proportion of voiceless-to-voiced consonants (ZOIB beta)
- (final) vowel quality (high, mid, low)
- Prosody (1vs0)
- Long vowel vs. short vowel (1vs0)

```{r}
# vowel vs no vowel

segs <- unique(str_split(paste0(v$form_add_all, collapse=""), "")[[1]])

# a decision made here: {j,w} in V+{j,w} is coded as vowel, but in {w,j}+v it is coded as consonant
vowels <- c("V","u","a",".w$","i","o","e","ə",".j$","ɯ","ẽ","õ","ɛ","ɔ","ê","ʊ","ɑ","æ","ã","ĩ","ɨ","y","A")
vowels.no.wj <- c("V","u","a","i","o","e","ə","ɯ","ẽ","õ","ɛ","ɔ","ê","ʊ","ɑ","æ","ã","ĩ","ɨ","y","A")
consonants <- c("n","t","w[VuaioeəɯẽõɛɔêʊɑæãĩɨyA]","m","h","j[VuaioeəɯẽõɛɔêʊɑæãĩɨyA]","r","s","l","k","ʔ","v","C","d","g","K","J","ŋ","p","q","χ")
obstr_vl <- c("t","h","s","k","ʔ","C","K","p","q","χ")

vowel_height <- list(
  `V`= NA,
  `u`= "h",
  `a`= "l",
  `i`= "h",
  `o`= "m",
  `e`= "m",
  `ə`= "m",
  `ɯ`= "h",
  `ẽ`= "m",
  `õ`= "m",
  `ɛ`= "m",
  `ɔ`= "m",
  `ê`= "m",
  `ʊ`= "h",
  `ɑ`= "l",
  `æ`= "l",
  `ã`= "l",
  `ĩ`= "h",
  `ɨ`= "h",
  `y`= "h",
  `A`= "l")

find_last_vowel <- function (str_vector, vowel_height, ignore.length=F) {
  if (ignore.length) {
    last_vowel <- unlist(
      map(
        str_extract_all(str_vector, paste0(names(vowel_height), collapse="|")), 
        function (x) ifelse(length(x) > 0, x[length(x)], NA)
      )
    )
  } else {
    last_vowel <- unlist(
      map(
        str_extract_all(str_vector, 
                        paste0(names(vowel_height), names(vowel_height), "|", names(vowel_height), "[jw]|", names(vowel_height), collapse="|")), 
        function (x) ifelse(length(x) > 0, x[length(x)], NA)
      )
    )
  }
  return(last_vowel)
}



find_height <- function (vowels, vowel_height) {
  return(vowel_height[vowels])
}

v <- v %>%
  mutate(
    # vowel vs no vowel
    vowel_pres = ifelse(form_add_all != "", grepl(paste0(vowels, collapse="|"),form_add_all), NA),
    # consonant vs no consonant
    cons_pres = ifelse(form_add_all != "", grepl(paste0(consonants, collapse="|"),form_add_all), NA),
    # does it end in a vowel or not?
    vowel_end = ifelse(form_add_all != "", grepl(paste0(paste0(vowels,"$"), collapse="|"),form_add_all), NA),
    # does it have a voiceless obstruent?
    obstr_vl_pres = ifelse(form_add_all != "", grepl(paste0(obstr_vl, collapse="|"),form_add_all), NA),
    # proportion of voiceless to voiced material
    obstr_vl_count = str_count(form_add_all, paste0(obstr_vl, collapse="|")),
    obstr_vl_prop = ifelse(form_add_all != "", obstr_vl_count / nchar(form_add_all), NA),
    obstr_vl_count = ifelse(form_add_all != "", obstr_vl_count, NA),
    # last V quality
    last_vowel = find_last_vowel(form_add_all, vowel_height),
    last_vowel_qual = find_last_vowel(form_add_all, vowel_height, ignore.length=T),
    v_height = as.vector(unlist(map(vowel_height[last_vowel_qual], function (x) ifelse(is.null(x), NA, x)))),
    # last V length
    v_long = ifelse(nchar(last_vowel)==2, T, F)
    )

```

```{r}

v$cons_end <- ifelse(v$cons_pres == TRUE & v$vowel_end == FALSE, TRUE, FALSE)
v$cons_onset <- ifelse(v$cons_pres == TRUE & v$cons_end == FALSE, TRUE, FALSE)

```


## Write data

```{r}
write_csv(v, "../derived_data/vocatives_processed.csv")
```
