---
title: "Publication ready plots"
project: "Vocatives"
author: "Marton Soskuthy & Timo Roettger"
date: "7/24/2019"
contacts: "timo.b.roettger@gmail.com, marton.soskuthy@ubc.ca"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Prep

Loading libraries & raw data.

```{r preprocess}

# load packages
library(tidyverse)
library(ggbeeswarm)
library(brms)

# load data
v <- read_csv('../derived_data/vocatives_processed.csv')
langs <- read_csv('../derived_data/voc_languages.csv')

# join them
v <- full_join(v,langs)

# store theme
theme_voc <- 
theme_classic() + 
  theme(legend.position = "none",
        legend.key.height = unit(2,"line"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.background = element_rect(fill = "transparent"),
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent"),
        strip.text.y = element_text(size = 12, hjust = 0),
        axis.text = element_text(size = 12),
        axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.margin = unit(c(0.4,0.4,0.4,0.4),"cm"))


```
```{r agg}

# aggregate raw data over lang family
v_aggr <- v %>%
    filter(case_broad != "nom-like") %>% 
    dplyr::select(language, case_broad, prosody_clean,
         vowel_pres, cons_pres, cons_end, vowel_end, 
         obstr_vl_pres, v_long, family1, latitude, longitude) %>% 
  group_by(language, case_broad, family1) %>%
  summarise(vowel_pres = mean(vowel_pres, na.rm = T),
            cons_pres = mean(cons_pres, na.rm = T),
            vowel_end = mean(vowel_end, na.rm = T),
            obstr_vl_pres = mean(obstr_vl_pres, na.rm = T),
            v_long = mean(v_long, na.rm = T),
            cons_end = mean(cons_end, na.rm = T),
            prosody_clean = mean(prosody_clean, na.rm = T),
            latitude = mean(latitude),
            longitude = mean(longitude)) %>%
  ungroup()

# aggregated over families
v_aggr_family <- v_aggr %>%
  group_by(family1, case_broad) %>%
  add_tally() %>% 
  filter(!is.na(case_broad)) %>% 
  summarise(vowel_pres = mean(vowel_pres, na.rm = T),
            cons_pres = mean(cons_pres, na.rm = T),
            vowel_end = mean(vowel_end, na.rm = T),
            obstr_vl_pres = mean(obstr_vl_pres, na.rm = T),
            v_long = mean(v_long, na.rm = T),
            cons_end = mean(cons_end, na.rm = T),
            prosody_clean = mean(prosody_clean, na.rm = T),
            n = mean(n)) %>%
  ungroup()

# aggregate raw data over lang family and suffix vs. non-suffix
v_aggr_suffix <- v %>%
    filter(case_broad != "nom-like") %>% 
    filter(!is.na(form_add_suffix), 
           morphological == "affix") %>% 
    dplyr::select(language, case_broad, prosody_clean,
         vowel_pres, cons_pres, cons_end, vowel_end, 
         obstr_vl_pres, v_long, family1, latitude, longitude) %>% 
  group_by(language, case_broad, family1) %>%
  summarise(vowel_pres = mean(vowel_pres, na.rm = T),
            cons_pres = mean(cons_pres, na.rm = T),
            vowel_end = mean(vowel_end, na.rm = T),
            obstr_vl_pres = mean(obstr_vl_pres, na.rm = T),
            v_long = mean(v_long, na.rm = T),
            cons_end = mean(cons_end, na.rm = T),
            prosody_clean = mean(prosody_clean, na.rm = T),
            latitude = mean(latitude),
            longitude = mean(longitude)) %>%
  ungroup()


# aggregated over families
v_aggr_family_suffix <- v_aggr_suffix %>%
  group_by(family1, case_broad) %>%
  add_tally() %>% 
  filter(!is.na(case_broad)) %>% 
  summarise(vowel_pres = mean(vowel_pres, na.rm = T),
            cons_pres = mean(cons_pres, na.rm = T),
            vowel_end = mean(vowel_end, na.rm = T),
            obstr_vl_pres = mean(obstr_vl_pres, na.rm = T),
            v_long = mean(v_long, na.rm = T),
            cons_end = mean(cons_end, na.rm = T),
            prosody_clean = mean(prosody_clean, na.rm = T),
            n = mean(n)) %>%
  ungroup()


v_aggr$case_broad <- as.factor(v_aggr$case_broad)
levels(v_aggr$case_broad) <- c("accusative-like", "vocative")

v_aggr_family$case_broad <- as.factor(v_aggr_family$case_broad)
levels(v_aggr_family$case_broad) <- c("accusative-like", "vocative")

v_aggr_suffix$case_broad <- as.factor(v_aggr_suffix$case_broad)
levels(v_aggr_suffix$case_broad) <- c("accusative-like", "vocative")

v_aggr_family_suffix$case_broad <- as.factor(v_aggr_family_suffix$case_broad)
levels(v_aggr_family_suffix$case_broad) <- c("accusative-like", "vocative")


v_acc_voc <- v %>%
  filter(case_broad != "nom-like") %>% 
  mutate(case_broad_sum=ifelse(case_broad=="vocative", 1, 0),
         case_broad_sum=case_broad_sum-mean(case_broad_sum))

```


## Prosodic Modulation

### Prepare

```{r prep prosody}

# load in model output
prosody_model <- readRDS('../models/mod_prosody_strict.rds')

# extract posteriors
mod_prosody_strict_psamp <- as.matrix(posterior_samples(prosody_model)[,1:2])

# accusative-like - all probabilities
mod_prosody_strict_acc_probs <- plogis(mod_prosody_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_prosody_strict_acc_probs), 5)
lower <- round(quantile(mod_prosody_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_prosody_strict_acc_probs, 0.975), 5)

prosody_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_prosody_strict_voc_probs <- plogis(mod_prosody_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_prosody_strict_voc_probs), 5)
lower <- round(quantile(mod_prosody_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_prosody_strict_voc_probs, 0.975), 5)

prosody_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
prosody_post <- rbind(prosody_acc_post, prosody_voc_post)
prosody_post$case_broad <- c("accusative-like", "vocative")


```

### Plot (publication plot Figure 2)

```{r plot prosody}

# plot and save
ggplot(data = prosody_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family, aes(x = case_broad, y = prosody_clean, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Prosodic manipulation",
       subtitle = "Vocatives are often prosodically marked\n",
       y = "proportion of forms \nwith prosodic changes\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/prosody.pdf", width = 4.5, height = 3.6)
ggsave("../plots/prosody.png", width = 4.5, height = 3.6)


```


## Consonants presence

### Prepare

```{r prep consonants}

# load in model output for overall consonant model
cons_pres_model <- readRDS('../models/mod_consonants.rds')

# extract posteriors
mod_cons_pres_strict_psamp <- as.matrix(posterior_samples(cons_pres_model)[,1:2])

# accusative-like - all probabilities
mod_cons_pres_strict_acc_probs <- plogis(mod_cons_pres_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_pres_strict_acc_probs), 5)
lower <- round(quantile(mod_cons_pres_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_pres_strict_acc_probs, 0.975), 5)

cons_pres_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_cons_pres_strict_voc_probs <- plogis(mod_cons_pres_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_pres_strict_voc_probs), 5)
lower <- round(quantile(mod_cons_pres_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_pres_strict_voc_probs, 0.975), 5)

cons_pres_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
cons_pres_post <- rbind(cons_pres_acc_post, cons_pres_voc_post)
cons_pres_post$case_broad <- c("accusative-like", "vocative")

#####

# load in model output for presence of consonants only in suffixes
cons_pres_suff_model <- readRDS('../models/mod_consonants_suff.rds')

# extract posteriors
mod_cons_pres_suff_strict_psamp <- as.matrix(posterior_samples(cons_pres_suff_model)[,1:2])

# accusative-like - all probabilities
mod_cons_pres_suff_strict_acc_probs <- plogis(mod_cons_pres_suff_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_pres_suff_strict_acc_probs), 5)
lower <- round(quantile(mod_cons_pres_suff_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_pres_suff_strict_acc_probs, 0.975), 5)

cons_pres_suff_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_cons_pres_suff_strict_voc_probs <- plogis(mod_cons_pres_suff_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_pres_suff_strict_voc_probs), 5)
lower <- round(quantile(mod_cons_pres_suff_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_pres_suff_strict_voc_probs, 0.975), 5)

cons_pres_suff_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
cons_pres_suff_post <- rbind(cons_pres_suff_acc_post, cons_pres_suff_voc_post)
cons_pres_suff_post$case_broad <- c("accusative-like", "vocative")    
  

```

### Plot overall pattern

```{r plot consonant presence}

# plot and save
ggplot(data = cons_pres_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family, aes(x = case_broad, y = cons_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Presence of consonant",
       subtitle = "Overall: Vocatives often have no consonants\n",
       y = "proportion of forms \nwith a consonant\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/consonant.pdf", width = 4.5, height = 3.6)
ggsave("../plots/consonant.png", width = 4.5, height = 3.6)


```

### Plot pattern for suffixes only

```{r plot consonant presence for suffixes only}

# plot and save
ggplot(data = cons_pres_suff_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family_suffix, aes(x = case_broad, y = cons_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Consonants in suffixes",
       subtitle = "Suffixes only: Vocatives often have no consonants\n",
       y = "proportion of forms \nwith a consonant\n", 
       x = "\n") +
  theme_voc


# save  
ggsave("../plots/consonant_suffix.pdf", width = 4.5, height = 3.6)
ggsave("../plots/consonant_suffix.png", width = 4.5, height = 3.6)


```

## Final consonant?

### Prepare

```{r prep consonants final}

# load in model output for overall consonant model
cons_final_model <- readRDS('../models/mod_c_final.rds')

# extract posteriors
mod_cons_final_strict_psamp <- as.matrix(posterior_samples(cons_final_model)[,1:2])

# accusative-like - all probabilities
mod_cons_final_strict_acc_probs <- plogis(mod_cons_final_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_final_strict_acc_probs), 5)
lower <- round(quantile(mod_cons_final_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_final_strict_acc_probs, 0.975), 5)

cons_final_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_cons_final_strict_voc_probs <- plogis(mod_cons_final_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_final_strict_voc_probs), 5)
lower <- round(quantile(mod_cons_final_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_final_strict_voc_probs, 0.975), 5)

cons_final_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
cons_final_post <- rbind(cons_final_acc_post, cons_final_voc_post)
cons_final_post$case_broad <- c("accusative-like", "vocative")

```

### Plot

```{r plot consonant final}

v_aggr_family_cons <- v_aggr_family %>% 
  filter(cons_pres == T)

# plot and save
ggplot(data = cons_final_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family_cons, aes(x = case_broad, y = cons_end, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Position of consonant",
       subtitle = "Vocative do not have less final consonants\n",
       y = "proportion of forms \nwith a final consonant\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/consonant_final.pdf", width = 4.5, height = 3.6)
ggsave("../plots/consonant_final.png", width = 4.5, height = 3.6)


```

## Voiceless consonant?

### Prepare

```{r prep consonants voiceless}

# load in model output for overall consonant model
cons_vl_model <- readRDS('../models/mod_vl_obstr.rds')

# extract posteriors
mod_cons_vl_strict_psamp <- as.matrix(posterior_samples(cons_vl_model)[,1:2])

# accusative-like - all probabilities
mod_cons_vl_strict_acc_probs <- plogis(mod_cons_vl_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_vl_strict_acc_probs), 5)
lower <- round(quantile(mod_cons_vl_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_vl_strict_acc_probs, 0.975), 5)

cons_vl_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_cons_vl_strict_voc_probs <- plogis(mod_cons_vl_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_cons_vl_strict_voc_probs), 5)
lower <- round(quantile(mod_cons_vl_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_cons_vl_strict_voc_probs, 0.975), 5)

cons_vl_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
cons_vl_post <- rbind(cons_vl_acc_post, cons_vl_voc_post)
cons_vl_post$case_broad <- c("accusative-like", "vocative")

```

### Plot 

```{r plot consonant voiceless}

v_aggr_family_cons <- v_aggr_family %>% 
  filter(cons_pres == T)

# plot and save
ggplot(data = cons_vl_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family_cons, aes(x = case_broad, y = obstr_vl_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Voicing of consonant",
       subtitle = "Vocative do not have less voiceless consonants\n",
       y = "proportion of forms \nwith voiceless consonants\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/consonant_vl.pdf", width = 4.5, height = 3.6)
ggsave("../plots/consonant_vl.png", width = 4.5, height = 3.6)


```

## Presence of vowel?

### Prepare

```{r prep vowel presence}

# load in model output for overall consonant model
vow_pres_model <- readRDS('../models/mod_vowel_pres.rds')

# extract posteriors
mod_vow_pres_strict_psamp <- as.matrix(posterior_samples(vow_pres_model)[,1:2])

# accusative-like - all probabilities
mod_vow_pres_strict_acc_probs <- plogis(mod_vow_pres_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_vow_pres_strict_acc_probs), 5)
lower <- round(quantile(mod_vow_pres_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_vow_pres_strict_acc_probs, 0.975), 5)

vow_pres_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_vow_pres_strict_voc_probs <- plogis(mod_vow_pres_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_vow_pres_strict_voc_probs), 5)
lower <- round(quantile(mod_vow_pres_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_vow_pres_strict_voc_probs, 0.975), 5)

vow_pres_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
vow_pres_post <- rbind(vow_pres_acc_post, vow_pres_voc_post)
vow_pres_post$case_broad <- c("accusative-like", "vocative")

```

### Plot

```{r plot vowel presence}

# plot and save
ggplot(data = vow_pres_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family, aes(x = case_broad, y = vowel_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Results: Presence of vowel",
       subtitle = "Most forms have a vowel\n",
       y = "proportion of forms \nwith vowels\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/vowels.pdf", width = 4.5, height = 3.6)
ggsave("../plots/vowels.png", width = 4.5, height = 3.6)


```


## Vowel quantity?

### Prepare

```{r prep vowel quant}

# load in model output for overall consonant model
vow_quant_model <- readRDS('../models/mod_v_long.rds')

# extract posteriors
mod_vow_quant_strict_psamp <- as.matrix(posterior_samples(vow_quant_model)[,1:2])

# accusative-like - all probabilities
mod_vow_quant_strict_acc_probs <- plogis(mod_vow_quant_strict_psamp %*% matrix(c(1, min(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_vow_quant_strict_acc_probs), 5)
lower <- round(quantile(mod_vow_quant_strict_acc_probs, 0.025), 5)
upper <- round(quantile(mod_vow_quant_strict_acc_probs, 0.975), 5)

vow_quant_acc_post <- data.frame(mean, lower, upper, row.names = NULL)

# vocative-like - all probabilities
mod_vow_quant_strict_voc_probs <- plogis(mod_vow_quant_strict_psamp %*% matrix(c(1, max(v_acc_voc$case_broad_sum))))

# get posterior mean and 95% CI for accusative
mean <- round(mean(mod_vow_quant_strict_voc_probs), 5)
lower <- round(quantile(mod_vow_quant_strict_voc_probs, 0.025), 5)
upper <- round(quantile(mod_vow_quant_strict_voc_probs, 0.975), 5)

vow_quant_voc_post <- data.frame(mean, lower, upper, row.names = NULL)

# merge
vow_quant_post <- rbind(vow_quant_acc_post, vow_quant_voc_post)
vow_quant_post$case_broad <- c("accusative-like", "vocative")

```

### Plot 

```{r plot vowel quant}

# plot and save
ggplot(data = vow_quant_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family, aes(x = case_broad, y = v_long, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  labs(title = "Results: Vowel length",
       subtitle = "Vocatives do not have more long vowels\n",
       y = "proportion of forms \nwith long vowels\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/vowels_long.pdf", width = 4.5, height = 3.6)
ggsave("../plots/vowels_long.png", width = 4.5, height = 3.6)


```

## Vowel quality?

### Prepare

```{r prep vowel qual}

# subset
v_acc_voc_v_height <- filter(v_acc_voc, !is.na(form_add_all) & vowel_pres & !is.na(v_height))

# raw data by language family
v_acc_voc_v_height_agg <- 
v_acc_voc_v_height %>%
  count(family1, case_broad, v_height) %>%
  complete(nesting(family1, case_broad), v_height, fill = list(n = 0)) %>% 
  group_by(family1, case_broad) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup()

v_acc_voc_v_height_agg$case_broad <- as.factor(v_acc_voc_v_height_agg$case_broad)
levels(v_acc_voc_v_height_agg$case_broad) <- c("accusative-like", "vocative")


# load in model output for overall consonant model
vow_quant_model <- readRDS('../models/mod_v_height.rds')

# extract posteriors
psamp <- posterior_samples(vow_quant_model)

acc_matrix <- matrix(c(1, min(v_acc_voc_v_height$case_broad_sum)))
voc_matrix <- matrix(c(1, max(v_acc_voc_v_height$case_broad_sum)))
l_over_h_samp <- as.matrix(dplyr::select(psamp, matches("b_mul")))
m_over_h_samp <- as.matrix(dplyr::select(psamp, matches("b_mum")))

# probabilities!
p_h_acc <- 1 / (1 + exp(l_over_h_samp %*% acc_matrix) + exp(m_over_h_samp %*% acc_matrix))
p_h_voc <- 1 / (1 + exp(l_over_h_samp %*% voc_matrix) + exp(m_over_h_samp %*% voc_matrix))
p_l_acc <- p_h_acc * exp(l_over_h_samp %*% acc_matrix)
p_l_voc <- p_h_voc * exp(l_over_h_samp %*% voc_matrix)
p_m_acc <- p_h_acc * exp(m_over_h_samp %*% acc_matrix)
p_m_voc <- p_h_voc * exp(m_over_h_samp %*% voc_matrix)

# means / quantiles / differences
# low vowels
low_acc_mean <- mean(p_l_acc)
low_acc_lower <- quantile(p_l_acc, 0.025)
low_acc_upper <- quantile(p_l_acc, 0.975)

low_acc_post <- data.frame(low_acc_mean, low_acc_lower, low_acc_upper, row.names = NULL)
colnames(low_acc_post) <- c("mean", "lower", "upper")

low_voc_mean <- mean(p_l_voc)
low_voc_lower <- quantile(p_l_voc, 0.025)
low_voc_upper <- quantile(p_l_voc, 0.975)

low_voc_post <- data.frame(low_voc_mean, low_voc_lower, low_voc_upper, row.names = NULL)
colnames(low_voc_post) <- c("mean", "lower", "upper")

# mid vowels
mid_acc_mean <- mean(p_m_acc)
mid_acc_lower <- quantile(p_m_acc, 0.025)
mid_acc_upper <- quantile(p_m_acc, 0.975)

mid_acc_post <- data.frame(mid_acc_mean, mid_acc_lower, mid_acc_upper, row.names = NULL)
colnames(mid_acc_post) <- c("mean", "lower", "upper")

mid_voc_mean <- mean(p_m_voc)
mid_voc_lower <- quantile(p_m_voc, 0.025)
mid_voc_upper <- quantile(p_m_voc, 0.975)

mid_voc_post <- data.frame(mid_voc_mean, mid_voc_lower, mid_voc_upper, row.names = NULL)
colnames(mid_voc_post) <- c("mean", "lower", "upper")

# high vowels
high_acc_mean <- mean(p_h_acc)
high_acc_lower <- quantile(p_h_acc, 0.025)
high_acc_upper <- quantile(p_h_acc, 0.975)

high_acc_post <- data.frame(high_acc_mean, high_acc_lower, high_acc_upper, row.names = NULL)
colnames(high_acc_post) <- c("mean", "lower", "upper")

high_voc_mean <- mean(p_h_voc)
high_voc_lower <- quantile(p_h_voc, 0.025)
high_voc_upper <- quantile(p_h_voc, 0.975)

high_voc_post <- data.frame(high_voc_mean, high_voc_lower, high_voc_upper, row.names = NULL)
colnames(high_voc_post) <- c("mean", "lower", "upper")

# merge
vow_qual_post <- rbind(low_acc_post, low_voc_post,
                       mid_acc_post, mid_voc_post,
                       high_acc_post, high_voc_post)

# add v_height and case_broad
vow_qual_post$case_broad <- rep(c("accusative-like", "vocative"), 3)
vow_qual_post$v_height <- rep(c("l", "m", "h"), each = 2)

```

### Plot (publication plot Figure 4)

```{r vowel qual plot}

# rename and relevel v_height
vow_qual_post$v_height <- as.factor(vow_qual_post$v_height)
v_acc_voc_v_height_agg$v_height <- as.factor(v_acc_voc_v_height_agg$v_height)

levels(vow_qual_post$v_height) <- c("high", "low", "mid")
levels(v_acc_voc_v_height_agg$v_height) <- c("high", "low", "mid")

# reorder factor levels
source("https://raw.githubusercontent.com/janhove/janhove.github.io/master/RCode/sortLvls.R")
vow_qual_post$v_height <- sortLvls.fnc(vow_qual_post$v_height, c(2,3,1))
v_acc_voc_v_height_agg$v_height <- sortLvls.fnc(v_acc_voc_v_height_agg$v_height, c(2,3,1))


# plot
ggplot(data = vow_qual_post, aes(x = case_broad, y = mean, fill = case_broad)) +
  geom_point(data = v_acc_voc_v_height_agg, aes(x = case_broad, y = prop, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  facet_grid(~ v_height) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  labs(title = "Results: Vowel quality",
       subtitle = "Vocatives have more mid vowels and fewer high vowels\n",
       y = "proportion of forms \nwith respective vowel quality\n", 
       x = "\n") +
  theme_voc

# save  
ggsave("../plots/vowels_qual.pdf", width = 8, height = 3.6)
ggsave("../plots/vowels_qual.png", width = 8, height = 3.6)


```

## Plot maps of the major finding

### Consonants

```{r map plots}

library(maps)

world_map <- map_data("world")

map <- ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "white", size = 0.2) +
  geom_point(data = v_aggr_suffix, aes(x = longitude, y = latitude, color = cons_pres, group = NA, size = 4),
             alpha = 0.8, position = position_jitter(0.5,0.5)) +
  coord_map(xlim = c(-180, 180),ylim = c(-60, 80)) +
  #coord_fixed(ylim = c(-60, 80), ratio = 1) +
  #ylim(-60,80) +
  facet_grid(~case_broad) +
  scale_colour_gradient(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1), low = "#4dac26", high = "#d01c8b") +
  labs(title = "Map: Presence of consonants in suffixes",
       subtitle = "Vocatives often have no consonants\n") +
  theme_voc +
  theme(
    legend.position = "right",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank())

map

# save  
ggsave("../plots/maps_consonants_suffixes.pdf", width = 10, height = 3.6)
ggsave("../plots/maps_consonants_suffixes.png", width = 10, height = 3.6)


```

### Combine plots (publication plot Figure 3)

```{r combine plots}

library(ggpubr)

# plot map
map_combo <- ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "white", size = 0.2) +
  geom_point(data = v_aggr, aes(x = longitude, y = latitude, color = cons_pres, group = NA), alpha = 0.8, size = 3, position = position_jitter(0.1)) +
   coord_map(xlim = c(-180, 180),ylim = c(-60, 80)) +
  #ylim(-60,80) +
  #coord_fixed(ylim = c(-60, 80), ratio = 1) +
  facet_grid(~case_broad) +
  scale_colour_gradient(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1), low = "#4dac26", high = "#d01c8b") +
  labs(title = "A - Map: Presence of consonants",
       subtitle = "Vocatives often have no consonants\n") +
  theme_voc +
  theme(
    legend.position = "right",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank())

# plot consonant overall
cons_overall_combo <- ggplot(data = cons_pres_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family, aes(x = case_broad, y = cons_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "B - Overall results",
       #subtitle = "Overall: Vocatives often have no consonants\n",
       y = "proportion of forms \nwith a consonant\n", 
       x = "\n") +
  theme_voc

# plot consonant suffix
cons_suffix_combo <- 
  ggplot(data = cons_pres_suff_post, aes(x = case_broad, y = mean)) +
  geom_point(data = v_aggr_family_suffix, aes(x = case_broad, y = cons_pres, 
                                       color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  #geom_line(aes( group = 1)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.075) +
  geom_point(aes(fill = case_broad), size = 6, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "C - Results for suffixes only",
       #subtitle = "Suffixes only: Vocatives often have no consonants\n",
       y = " \n", 
       x = "\n") +
  theme_voc

map_and_point <- 
  ggarrange(map_combo, hjust = -1,
            ggarrange(cons_overall_combo, cons_suffix_combo,
                      ncol = 2, 
                      #labels = c("B", "C"), font.label = list(size = 14),
                      widths = c(1,1)),
            heights = c(1,1),
            #labels = "A", font.label = list(size = 14),
            nrow = 2,
            common.legend = F)

map_and_point

# save  
ggsave("../plots/maps_combined.pdf", width = 9, height = 7)
ggsave("../plots/maps_combined.png", width = 9, height = 7)


```



