---
title: "First pass analysis of vocatives data"
author: "Márton Sóskuthy"
date: "18/06/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

Loading libraries & raw data.

```{r}
library(tidyverse)
library(ggbeeswarm)

v <- read_csv('../data/vocatives_processed.csv')
langs <- read_csv('../derived_data/voc_languages.csv')

v <- full_join(v,langs)

v_aggr <- v %>%
  dplyr::select(language, case_broad, prosody_clean,
         vowel_pres,cons_pres,cons_end,vowel_end,obstr_vl_pres,
         v_long, family1, latitude, longitude) %>% 
  group_by(language, case_broad, family1) %>%
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long = mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T),
            latitude=mean(latitude),
            longitude=mean(longitude)) %>%
  ungroup()

v_agg_overall <- v_aggr %>%
  group_by(case_broad) %>%
  filter(!is.na(case_broad)) %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long=mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T)) %>%
  ungroup()


# get aggregates for largest families
v_aggr_family <- v_aggr %>%
  group_by(family1, case_broad) %>%
      add_tally() %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long=mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T),
            n = mean(n)) %>%
  ungroup()




# only affixes with consonants

v_aggr_c <- v %>%
  dplyr::select(language, case_broad, prosody_clean,
         vowel_pres,cons_pres,cons_end,vowel_end,obstr_vl_pres,
         v_long, family1) %>% 
  filter(cons_pres) %>%
  group_by(language, case_broad, family1) %>%
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T)) %>%
  ungroup()

v_agg_overall_c <- v_aggr_c %>%
  group_by(case_broad) %>%
  filter(!is.na(case_broad)) %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T)) %>%
  ungroup()

v_aggr_family_c <- v_aggr_c %>%
  group_by(family1, case_broad) %>%
      add_tally() %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            n = mean(n)) %>%
  ungroup()


## aggregates for suffixes

v_aggr_suffix <- v %>%
  filter(!is.na(form_add_suffix), morphological=="affix") %>%
  dplyr::select(language, case_broad, prosody_clean,
         vowel_pres,cons_pres,cons_end,vowel_end,obstr_vl_pres,
         v_long, family1, latitude, longitude) %>% 
  group_by(language, case_broad, family1) %>%
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long = mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T),
            latitude=mean(latitude),
            longitude=mean(longitude)) %>%
  ungroup()

v_agg_overall_suffix <- v_aggr_suffix %>%
  group_by(case_broad) %>%
  filter(!is.na(case_broad)) %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long=mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T)) %>%
  ungroup()


# get aggregates for largest families
v_aggr_family_suffix <- v_aggr_suffix %>%
  group_by(family1, case_broad) %>%
      add_tally() %>% 
  summarise(vowel_pres=mean(vowel_pres, na.rm=T),
            cons_pres=mean(cons_pres, na.rm=T),
            vowel_end=mean(vowel_end, na.rm=T),
            obstr_vl_pres=mean(obstr_vl_pres, na.rm=T),
            v_long=mean(v_long, na.rm=T),
            cons_end = mean(cons_end, na.rm=T),
            prosody_clean=mean(prosody_clean, na.rm=T),
            n = mean(n)) %>%
  ungroup()



#store theme

theme_voc <- 
theme_classic() + 
  theme(legend.position = "none",
        legend.key.height = unit(2,"line"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.background = element_rect(fill = "transparent"),
        strip.background = element_blank(),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent"),
        strip.text.y = element_text(size = 12, hjust = 0),
        axis.text = element_text(size = 12),
        axis.line = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.margin = unit(c(0.3,0.3,0.3,0.3),"cm"))

```

## Aggregating & plotting

Case morphemes generally have vowels, with a few exceptions (especially nominatives). No real vocative advantage.

```{r}

v_aggr_family %>%
  filter(!is.nan(vowel_pres), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(vowel_pres = mean(vowel_pres),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = vowel_pres)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",],
                aes(ymin = vowel_pres + 0.1, ymax = vowel_pres - 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Vocatives contain vowels as frequently as other cases",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with vowels", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/vowel_pres.pdf", width=6, height=4.5)
```


Vocatives have less consonants.

```{r}

v_aggr_family %>%
  filter(!is.nan(cons_pres), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(cons_pres = mean(cons_pres),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = cons_pres)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",],
                aes(ymin = cons_pres + 0.1, ymax = cons_pres - 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Consonants less frequent in vocative markers",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with consonants", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/cons_pres.pdf", width=6, height=4.5)

```

Same but just with suffixes? Oh yeah.

```{r}
v_aggr_family_suffix %>%
  filter(!is.nan(cons_pres), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(cons_pres = mean(cons_pres),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = cons_pres)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall_suffix[v_agg_overall_suffix$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall_suffix[v_agg_overall_suffix$case_broad != "nom-like",],
                aes(ymin = cons_pres + 0.1, ymax = cons_pres - 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall_suffix[v_agg_overall_suffix$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Consonants less frequent in vocative suffixes",
       subtitle = "semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with consonants", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/cons_pres_suffix.pdf", width=6, height=4.5)

```

There is a vocative advantage for V-final: about 12% more final vowels for vocatives... (! when there is a consonant - otherwise this is heavily confounded with presence of consonant) MEH

```{r}

v_aggr_family_c %>%
  filter(!is.nan(vowel_end), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(vowel_end = mean(vowel_end),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = vowel_end)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall_c[v_agg_overall_c$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall_c[v_agg_overall_c$case_broad != "nom-like",],
                aes(ymin = vowel_end + 0.1, ymax = vowel_end - 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall_c[v_agg_overall_c$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Vocatives and other cases end in Vs equally frequently",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with final vowels", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/vowel_final_when_c_pres.pdf", width=6, height=4.5)
```


Vocatives have more long vowels!

```{r}

v_aggr_family %>%
  filter(!is.nan(v_long), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(v_long = mean(v_long),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = v_long)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",],
                aes(ymin = v_long + 0.1, ymax = v_long - 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Vocatives have more long vowels",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with long vowels", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/v_long.pdf", width=6, height=4.5)
```

Prosody -- was there any doubt about this one???

```{r}

v_aggr_family %>%
  filter(!is.nan(prosody_clean), 
         case_broad != "nom-like") %>%
  group_by(case_broad, family1) %>%
  summarise(prosody_clean = mean(prosody_clean),
            n = mean(n)) %>%
  ggplot(aes(x = case_broad, y = prosody_clean)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = n), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",],
                aes(ymin = prosody_clean - 0.1, ymax = prosody_clean + 0.1), width = 0.1
                ) +
  geom_point(data = v_agg_overall[v_agg_overall$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  labs(title = "Vocatives are often prosodically marked",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of forms with prosodic changes", 
       x = "\n") +
  theme_voc
ggsave("../prelim_plots/prosody.pdf", width=6, height=4.5)
```

Less high vowels, and more mid-vowels in vocatives

```{r}

#
v_quali_aggr <- v %>%
  dplyr::select(language, case_broad, v_height, family1) %>% 
  filter(!is.na(v_height),
         case_broad != "nom-like") %>%
  group_by(language, case_broad, family1, v_height) %>%
  tally() %>%
  complete(nesting(language, family1, case_broad), v_height, fill = list(n = 0)) %>%
  group_by(language, family1, case_broad) %>%
  mutate(prop = n/sum(n),
         sum = sum(n)) %>%
  ungroup()
 
v_quali_aggr_family <- v %>%
  dplyr::select(language, case_broad, v_height, family1) %>% 
  filter(!is.na(v_height),
         case_broad != "nom-like") %>%
  group_by(case_broad, family1, v_height) %>%
  tally() %>%
  complete(nesting(family1, case_broad), v_height, fill = list(n = 0)) %>%
  group_by(family1, case_broad) %>%
  mutate(prop = n/sum(n),
         sum = sum(n)) %>%
  ungroup()


v_quali_aggr_overall <- v %>%
  dplyr::select(language, case_broad, v_height, family1) %>% 
  filter(!is.na(v_height),
         case_broad != "nom-like") %>%
  group_by(case_broad, v_height) %>%
  tally() %>%
  complete(nesting(case_broad), v_height, fill = list(n = 0)) %>%
  group_by(case_broad) %>%
  mutate(prop = n/sum(n),
         sum = sum(n)) %>%
  ungroup()

```

```{r}

v_quali_aggr_family %>% 
  ggplot(aes(x = case_broad, y = prop)) +
  #geom_quasirandom(aes(color = case_broad, size = n), alpha = 0.3) +
  geom_point(aes(color = case_broad, size = sum), alpha = 0.3,
             position = position_quasirandom(width = 0.2)) +
  geom_line(data = v_quali_aggr_overall[v_quali_aggr_overall$case_broad != "nom-like",], 
             aes(group = 1)) +
  geom_errorbar(data = v_quali_aggr_overall[v_quali_aggr_overall$case_broad != "nom-like",],
                aes(ymin = prop + 0.1, ymax = prop - 0.1), width = 0.1
                ) +
  geom_point(data = v_quali_aggr_overall[v_quali_aggr_overall$case_broad != "nom-like",], 
             aes(fill = case_broad), size = 8, shape = 21) +
  scale_fill_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_color_manual(guide = guide_legend(title = "Case"),
                    values = c("#f1a340", "#998ec3")) +
  scale_size_continuous(guide = FALSE, range =c(2,6)) +
  scale_y_continuous(breaks = (c(0,0.25,0.5,0.75,1)), limits = c(0,1)) + 
  facet_grid(~factor(v_height, levels=c("l","m","h"))) +
  labs(title = "Vocatives typically contain mid vowels",
       subtitle = "Semi-transparent points are averages for language families\n",
       y = "\nproportion of occurences", 
       x = "\n") +
  theme_voc +
  theme(strip.text = element_text(size = 12))
ggsave("../prelim_plots/v_height.pdf", width=6, height=4.5)
```





Old stuff

```{r}
v_aggr_qual <- v %>%
  filter(!is.na(v_height),
         case_broad != "nom-like") %>%
   group_by(language, case_broad, v_height, family1) %>%
  #        add_tally() %>% 
  group_by(language, case_broad, v_height, family1) %>%
  tally() %>%
  ungroup() %>%
  complete(nesting(language, case_broad), v_height, fill=list(n = 0)) %>%
  group_by(language, case_broad) %>%
  mutate(prop=n/sum(n),
         sum = sum(n)) %>%
  ungroup()

v_aggr_qual %>%
  group_by(case_broad, v_height) %>%
  summarise(prop=mean(prop)) %>%
  ggplot(aes(x=factor(v_height, levels=c("l","m","h")), y=prop, fill = v_height)) + 
  facet_grid(~case_broad) +
  geom_point(stat="identity")

```

## Further exploration

```{r}
v %>% count(case_broad, morphological)

```

## An attempt at modelling the data

Let's look at our massive presence-of-consonants effect.

```{r}
v_cons_pres <- filter(v, case_broad != "nom-like" & !is.na(form_add) & form_add != "zero")

v_cons_pres$case_broad_f <- as.factor(v_cons_pres$case_broad)
v_cons_pres$family1_f <- as.factor(v_cons_pres$family1)

mod_cons_pres <- glmer(cons_pres ~ case_broad +  (1 + case_broad | family1),
             data=v_cons_pres, family="binomial")
mod_cons_pres_b <- brm(cons_pres ~ case_broad + (1 + case_broad | family1) + s(latitude,longitude,bs="sos",k=10, by=case_broad),
            data=v_cons_pres, family="bernoulli", cores=2, control=list(adapt_delta=0.99))


mod_cons_pres_gam <- bam(cons_pres ~ case_broad_f + s(family1_f, bs="re") + s(family1_f, case_broad_f, bs="re") + s(latitude,longitude,bs="sos",k=10, by=case_broad_f),
            data=v_cons_pres, family="binomial", discrete=T)

fvisgam(mod_cons_pres_gam, view=c("longitude","latitude"), cond=list(case_broad_f="acc-like"))
map('world', add=T)
fvisgam(mod_cons_pres_gam, view=c("longitude","latitude"), cond=list(case_broad_f="vocative"))
map('world', add=T)



summary(mod_cons_pres)
summary(mod_cons_pres_b)
summary(mod_cons_pres_gam)

ms <- marginal_smooths(mod_cons_pres_b)[[1]]


ggplot(ms, aes(x=longitude, y=latitude)) +
  facet_grid(~case_broad) +
  geom_raster(aes(fill=estimate__)) +
  borders("world", xlim=range(ms$longitude), ylim=range(ms$latitude), col="black", lwd=0.2)

ggplot(filter(v_aggr, case_broad %in% c("acc-like", "vocative")), aes(x=longitude, y=latitude, col=cons_pres)) +
  facet_grid(~case_broad) +
  geom_point(size=3) +
  scale_colour_viridis_c() +
  #geom_raster(aes(fill=estimate__)) +
  borders("world", xlim=range(ms$longitude), ylim=range(ms$latitude), col="black", lwd=0.2) + theme_bw()



v_v_long <- filter(v, case_broad != "nom-like" & !is.na(form_add) & form_add != "zero" & !is.na(v_long))
mod_v_long <- glmer(v_long ~ case_broad +  (1 + case_broad | family1),
             data=v_v_long, family="binomial")
mod_v_long_b <- brm(v_long ~ case_broad + (1+case_broad | family1),
            data=v_v_long, family="bernoulli", cores=2, control=list(adapt_delta=0.99))
summary(mod_v_long)
summary(mod_v_long_b)


```

Logistic models with vowel height.

```{r}
v_v_long$mid <- v_v_long$v_height=="m"
mid.glm <- glmer(mid ~ case_broad + (1 + case_broad | family1),
                 data=v_v_long, family="binomial")
summary(mid.glm)

v_v_long$low <- v_v_long$v_height=="l"
low.glm <- glmer(low ~ case_broad + (1 + case_broad | family1),
                 data=v_v_long, family="binomial")
summary(low.glm)

v_v_long$hi <- v_v_long$v_height=="h"
hi.glm <- glmer(hi ~ case_broad + (1 + case_broad || family1),
                 data=v_v_long, family="binomial", control=glmerControl(optimizer="bobyqa"))
summary(hi.glm)
```